\chapter{Introduction}
\label{ch:introdution}
This project is about Deliveroo, a competitive multi-agent game played on a tile grid.
Agents move across the map, pick up parcels and carry them to delivery tiles to score points.
Parcel rewards decay over time, so the goal is to maximize total delivered score by collecting
and dropping off parcels efficiently before they lose value. 
Extra points can be earned by answering questions
and following instructions sent by the Admin in the game chat.
The agents of each team cooperate with each other to compete against other teams.

The aim is to develop two cooperating agents:
\begin{itemize}
\item {Agent 1, BDI physical executor}: performs the actual moving, picking, and delivering.
\item {Agent 2, LLM cognitive coordinator}: can carry the same execution as agent one, but needs to have the ability to interpret 
Admin chat messages, deriving tasks, rules and questions.
\end{itemize}
PDDL planning is used as an additional tool to solve complex situations.

All the code and the instuctions to run the agents can be found in the GitHub repository:\\ 
\url{https://github.com/ale-bena/asa-autobots.git}

\section{Initial state}
\label{sec:init_state}
Both agents connect to the socket via DjsConnect() and they receive the initial state of the game.
This last consists in the static grid map, containing width, height and the tiles, which are parsed via an enum:

\noindent
{\small
\begin{minipage}[t]{0.5\textwidth}
\begin{itemize}
    \item WALL
    \item SPAWN
    \item DELIVERY
\end{itemize}
\end{minipage}%
\begin{minipage}[t]{0.5\textwidth}
\begin{itemize}
    \item PAVEMENT
    \item CRATE\_SPAWN, CRATE
    \item ARROW\_UP, ARROW\_RIGHT,\\ARROW\_DOWN, ARROW\_LEFT
\end{itemize}
\end{minipage}
}

Each agent also recieves their own identity and state (id, name, x, y, score) \texttt{via onYou}, stored in beliefs.me.
The game parameters, such as observation distance, player capacity, parcel decay interval,
movement duration are received separately via onConfig and are kept in beliefs.config and used for calibration.

Before the BDI executor begins acting, the two teammates perform a startup handshake to guarantee
both are online. Each agent's execution loop is gated on a \texttt{synced} flag: while unsynchronized,
it periodically sends a \texttt{SYNC\_REQ} to its peer; upon receiving one, an agent replies with a
\texttt{SYNC\_ACK} and releases the loop.